### 2014-05-12 paper/paper.tex
Add supervisor listing and discuss it

\documentclass[10pt,letter]{article}

\usepackage{amsmath}
\usepackage{amssymb}
% packages that allow mathematical formatting

\usepackage{graphicx}
% package that allows you to include graphics

\usepackage{setspace}
% package that allows you to change spacing

\usepackage{algpseudocode}
\usepackage{algorithmicx}
\usepackage{algorithm}

\usepackage[sc,osf]{mathpazo}
\linespread{1.05}         % Palatino needs more leading
\usepackage[T1]{fontenc}

\fontencoding{T1}
\fontfamily{ppl}
\fontseries{m}
\fontshape{n}
\fontsize{10}{13}
% Set font size. The first parameter is the font size to switch to; the second
% is the \baselineskip to use. The unit of both parameters defaults to pt. A
% rule of thumb is that the baselineskip should be 1.2 times the font size.
\selectfont

\usepackage{parskip}
% Use the style of having no indentation with a space between paragraphs.

\usepackage{enumitem}
% Resume an enumerated list, continuing the old numbering, after some
% intervening text.

% \usepackage{fullpage}
% package that specifies normal margins

\usepackage{titling}
% move the title up a bit, wastes less space
\setlength{\droptitle}{-5em}

\usepackage[usenames,dvipsnames]{color}
% for \color used in listings
\usepackage{textcomp}
% for upquote used in listings
\usepackage{listings}
\lstset{
    breakatwhitespace, % somehow prevents lstinline from line splitting
    tabsize=2,
    rulecolor=,
    basicstyle=\footnotesize,
    upquote=true,
    aboveskip={1.5\baselineskip},
    columns=fixed,
    extendedchars=true,
    breaklines=true,
    prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
    frame=single,
    showtabs=false,
    showspaces=false,
    showstringspaces=false,
    keywordstyle=\color{Purple},
    identifierstyle=\color{Black},
    commentstyle=\color{BrickRed},
    stringstyle=\color{RubineRed},
    captionpos=b,
}
\lstloadlanguages{erlang}
\lstset{language=erlang}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Macros

\newcommand{\chubby}[0]{\textsc{Chubby}}
\newcommand{\phat}[0]{\textsc{Phat}}
\newcommand{\phatraid}[0]{\textsc{PhatRaid}}
\newcommand{\raid}[1]{\textsc{RAID #1}}
\newcommand{\paxos}[0]{\textsc{Paxos}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{PHAT RAID, yo}
\author{Andrew Johnson \and Daniel King \and Lucas Waye \and Scott Moore}
\date{May 13, 2014}

\maketitle

Distributed file stores, like Google's \chubby{} and CS260r's \phat{}, send all
files through a single master node which consistently replicates the file on
many slaves. In this scheme, the master node's throughput is a bottleneck for
storing large files. We present \phatraid{} which mitigates this bottleneck by
partitioning files across many \paxos{} groups which form a \raid{1} group.

\section{Introduction}

\section{Implementation}

\phatraid{} is implemented in Erlang, reusing our previous implementation of
\phat{}. The implementation is divided into three domains: the server cluster,
the client, and the raid-client. The server code implements a VR-like consensus
algorithm as well as simple file system. The client code exposes a file system
API which hides the communication bookkeeping. The raid-client provides a file
system API which transparently partitions files and reassembles them as they are
sent and received from the server cluster.

In Erlang, the basic unit of concurrency is a \emph{process} and an Erlang VM is
known as a \emph{node}. A process may communicate with other processes on its
node as well as processes on other nodes. Multiple nodes may be on the same
machine or distributed across many machines.

The server cluster is implemented as a collection of nodes running the same
server code. One of these nodes is designated the master node. The other nodes
are known as replicas.

The master and each replica has four components: the supervisor, the
(unfortunately named) server, VR, and the file system.

\begin{itemize}
\item the supervisor -- if any of the other three processes dies, it shuts all
  of them down and restarts all of them
\item the server -- if the node is the master, it passes messages on to VR; if
  the node is a replica, it redirects clients to the master
\item VR -- maintains a log of opaque messages, achieving consensus with other
  nodes in the cluster via the Viewstamped Replication protocol
\item the file system -- a simple hierarchical file system with locks
\end{itemize}

Each of these are implemented using Erlang behaviors. An Erlang behavior is a
framework which implements common patterns like servers and finite state
machines. In particular, the supervisor uses the \texttt{supervisor} behavior,
the server and file system both use the \texttt{gen\_server} behavior, and VR
uses the \texttt{gen\_fsm} behavior.

\begin{itemize}
\item The \texttt{supervisor} behavior provides a framework for restarting
  failed processes.
\item The \texttt{gen\_server} behavior provides a framework for many-client
  single-server interactions. We implemented functions to process messages
  formatted as Erlang tuples. The \texttt{gen\_server} behavior handles
  low-level socket interaction.
\item The \texttt{gen\_fsm} behavior generalizes the \texttt{gen\_server}
  behavior by permitting the server to have a finite number of states. Each
  state has a set of functions with which to process messages. This behavior can
  be seen as an ad hoc polymorphic variant of the \texttt{gen\_server} behavior.
\end{itemize}

Additionally, the \texttt{gen\_server} and \texttt{gen\_fsm} behaviors allow
the programmer to specify an arbitrary state value which will be passed around
\`{a} la functional reactive programming.

\lstinputlisting[firstline=6,lastline=18,float,
                 caption=Supervisor Behavior --- \texttt{phat.erl},label=supervisor,
                 numbers=left, firstnumber=6]
                {../phat.erl}

Listing \ref{supervisor} is taken from our supervisor code. When a supervisor is
started, the \texttt{supervisor} behavior looks for a unary procedure called
\lstinline!init!. The \lstinline!init! procedure returns a child specification which
is a pair of a restart specification and a list of children. In Listing
\ref{supervisor}, the restart specification, \lstinline!{one_for_all,1,5}!
specifies that if any one node dies, all nodes should be restarted, unless more
than one failure has occurred in the past five seconds. If more than one failure
occurs in five seconds, the supervisor kills all the children and shuts itself
down as well.

Each child in the list of children is a 6-tuple consisting of: child name,
initial procedure call, child transience, shutdown style, child type, necessary
modules.

\begin{itemize}
\item The initial procedure call is a 3-tuple of module, procedure name and
  argument list.
\item Every child's transience is specified as \lstinline!permanent!, meaning
  they should be restarted.
\item Every child has shutdown style \lstinline!1! meaning they will first be
  asked to terminate and one second later they will be forcibly terminated.
\item All of our children are workers, which are distinguished from
  sub-supervisors.
\item The necessary modules are dynamically loaded when the child processes is
  initialized, there is a one-to-one mapping between our children and code
  modules.
\end{itemize}

\section{Performance}

\section{Conclusion}

\end{document}